\documentclass{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathtools}
\usepackage{graphicx}
\usepackage{enumitem}
\usepackage[margin=1in]{geometry}
\usepackage{titling}

\DeclareMathOperator{\Var}{Var}
\renewcommand{\vec}[1]{\mathbf{#1}}

\setlength{\droptitle}{-7em}   % This is your set screw

\begin{document}

\title{Math Statistics\\ Semiweekly HW 14}
\author{Ozaner Hansha}
\date{November 7, 2020}
\maketitle

\subsection*{Question 1}
\noindent\textbf{Problem:} Suppose that, without a vaccine, a person in a given population has a $p_{\text{w/o}}=.1$ chance of contracting the flu in a given year. We want to test that, given a particular vaccine, a person has only a $p_{\text{w}}=.06$ chance of contracting the flu.

Suppose we have a sample size of $n=10$. In this case we have the following null hypothesis $H_0$, alternative hypothesis $H_1$, and condition for accepting the null hypothesis $\hat H_0$:
\begin{align*}
    H_0&\equiv p_w=.1\tag{prob. of infection w/ vaccine is unchanged}\\
    H_1&\equiv p_w=.06\tag{prob. of infection w/ vaccine is $.06$}\\
    \hat H_0&\equiv X\ge1\tag{At least 1 subject infected within a year}
    % \neg \hat H_0= \hat H_1&\equiv X=0\tag{No subject is infected within a year}
\end{align*}

Where $X$ is the number of people infected in our sample. Also let $X_{n,p}\sim B(n,p)$. Note that with a sample size of $n$ and prob $p_w$ of infection we have $X\sim X_{n,p_w}$.
\bigskip

Given these hypotheses and conditions for accepting them, what is the probability of a type I error? What about a type II error?
\bigskip

\noindent\textbf{Solution:} For a type I error, i.e. accepting a false alternative hypothesis, we have:
\begin{align*}
    P(\text{Type I error})&=P(\neg\hat H_0\mid H_0)\tag{def. of type I error}\\
    &=P(X=0\mid p_w=.1)\tag{from def. of hypotheses}\\
    &=P(X_{10,.1}=0)\tag{$X\sim X_{n,p_w}$}\\
    &=\binom{10}{0}(.1)^0(1-.1)^{10-0}\tag{pmf of binomial RV}\\
    &=.9^{10}\approx.348678
\end{align*}

And for a type II error, i.e. accepting a false null hypothesis, we have:
\begin{align*}
    P(\text{Type II error})&=P(\hat H_0\mid H_1)\tag{def. of type II error}\\
    &=P(X\ge0\mid p_w=.06)\tag{from def. of hypotheses}\\
    &=\sum_{i=1}^{10}P(X_{10,.06}=i)\tag{$X\sim X_{n,p_w}$}\\
    &=1-P(X_{10,.06}=0)\tag{complement}\\
    &=1-\binom{10}{0}(.06)^0(1-.06)^{10-0}\tag{pmf of binomial RV}\\
    &=1-(.94)^{10}\approx.461385
\end{align*}
\end{document}