\documentclass{article}
\usepackage{amsmath}
\usepackage{amssymb}

\begin{document}

\title{Intro to Math Reasoning HW 10b}
\author{Ozaner Hansha}
\date{November 7, 2018}
\maketitle

\section*{Problem 1}
\textbf{Problem:} Prove that for any integer $j\ge 1$:
$$j<10^{j+1}$$
\textbf{Solution:} First note that when $j=1$ we have:
$$1<10^{1+1}=1<100$$

Which is true. Now we just need to show the inductive step:
\begin{align*}
  j&<10^{j+1}\\
  10j&<10^{j+1}10\\
  10j&<10^{j+2}\\
\end{align*}

Now we just need to show that $j+1<10j$ for all $j\ge 1$. Now we will prove the statement $1<9j$ for any positive $j$. For the $j=1$ case we have $1<9$ and for the inductive step:
\begin{align*}
  1<9j\tag{given}\\
  9j<9j+9\\
  9j<9(j+1)\\
  1<9j<9(j+1)\\
  1<9(j+1)
\end{align*}

Now we can use this to prove the statement we set out to show:
\begin{align*}
  1<9j\tag{for any $j\ge1$}\\
  1+j<9j+j\\
  j+1<10j
\end{align*}

Now we can write them together as:
$$j+1<10j<10^{j+2}\implies j+1<10^{j+2}$$

And so by induction the statement is true.

% \section*{Problem 2}
% \subsection*{Part a}
% \textbf{Problem:}
% \\\\
% \textbf{Solution:}
%
% \subsection*{Part b}
% \textbf{Problem:}
% \\\\
% \textbf{Solution:}

\section*{Problem 3}
\textbf{Problem:} Prove that if $m\le0$ then $10^m-1$ is divisible by 9.
\\\\
\textbf{Solution:} First we show the base case of $m=1$:
$$10^1-1=9$$

And 9 is clearly divisible by itself. Now we show the inductive step:
\begin{align*}
  10^m-1&=9q\\
  10(10^m-1)&=10(9q)\\
  10^m10-10&=90q\\
  10^{m+1}-10&=90q\\
  10^{m+1}-10+9&=90q+9\\
  10^{m+1}-1&=9(10q+1)
\end{align*}

Where $10q+1\in\mathbb Z$ since $q$ is also an integer. And so we are done.

\section*{Problem 4}
\textbf{Problem:} Prove that am integer greater than or equal to 1 is divisible by 9 iff the sum of its digits are divisible by 9.
\\\\
\textbf{Solution:} (I'm starting the indexing at 0 and restating above result in terms of modulo) Note that the decimal expansion of a number $n$ is:
$$n=10^nd_n+10^{n-1}d_{n-1}\cdots+10d_1+d_0$$

And since $10\equiv 1\ (\operatorname{mod } 9)$ we have $10^m\equiv 1\ (\operatorname{mod } 9)$. This means in modulo arithmetic we can say:
$$n\equiv 1d_n+1d_{n-1}\cdots+1d_1+d_0\ (\operatorname{mod } 9)$$

Which is just the sum of its digits. And so if $n$ is divisible by 9 then the right hand side (the sum of the digits) must also be divisible by 9. This is because they are in the same equivalence class under mod 9.

\section*{Problem 5}
% \subsection*{Part a}
\textbf{Problem:} Prove that $j\ge 1$ and $b\ge 2$ imply that $j<b^{j+1}$
\\\\
\textbf{Solution:} We can prove this using a base case and 2 inductive steps. The base case is simply $1<2^2$ which is clearly true. Now we need to show that $j+1<b^{j+2}$ follows from $j<b^{j+1}$:
\begin{align*}
  j<b^{j+1}\\
  j<b^{j+2}\tag{multiply by positive integer, always bigger}\\
  j+1<b^{j+2}\tag{as long as $j\le b$}
\end{align*}

Now we need to prove $j<(b+1)^{j+1}$ which should be clear since $b+1>b$ and all the other variables are positive integers and so $j<b^{j+1}<(b+1)^{j+1}$.

And so using induction and the base case the statement is true for all $b\le 2$ and $j\le 1$

% \subsection*{Part b}
% \textbf{Problem:}
% \\\\
% \textbf{Solution:}
%
% \subsection*{Part c}
% \textbf{Problem:}
% \\\\
% \textbf{Solution:}

\end{document}
